---
title: "Data"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
library(ggplot2)  
library(tidyverse)  
library(lubridate)

knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  echo = T,
  warning = F,
  cache = F
)

```

## Data sources

* Crime data of New York City is downloaded from NYC Open Data :
[NYPD Complaint Data Historic](https://data.cityofnewyork.us/Public-Safety/NYPD-Complaint-Data-Historic/qgea-i56i).

This dataset contains 780 million felony, misdemeanor, and violation crimes reported to the New York City Police Department (NYPD) from 1910 to 2021. Each record contains detailed information of a single crime reported to NYPD. In our project, we select the records from 2016-01-01 to 2017-12-31 to do analysis and modeling.

* Weather data is obtained from website of [NOAA](https://www.ncei.noaa.gov/metadata/geoportal/rest/metadata/item/gov.noaa.ncdc:C00861/html)


```{r}
crime_raw <- read.csv("data/nyc_complaint_data.csv")
crime_df <- crime_raw %>%
  janitor::clean_names() %>%
  mutate(
    date = mdy(cmplnt_fr_dt),
    hour_of_day = as.POSIXct(cmplnt_fr_tm, format="%H:%M:%OS"),
    borough = boro_nm,
    level = law_cat_cd,
    offense = ofns_desc,
    offense_classification = pd_desc,
    place = prem_typ_desc,
    suspect_age = susp_age_group,
    suspect_race = factor(susp_race),
    suspect_sex = factor(recode(susp_sex, "U" = "Unkonwn", "F" = "Female", "M" = "Male")),
    victim_age = vic_age_group,
    victim_race = factor(vic_race),
    victim_sex = factor(recode(vic_sex, "F" = "Female", "M" = "Male", "D" = "Business/Organization", "E" = "PSNY/People of the State of New York"))) %>%
  mutate(
    day_of_week = factor(weekdays(date)),
    mon = month(date),
    hour_of_day = hour(hour_of_day)
  )


crime_df <- crime_df %>%
 select(date:mon, latitude,longitude)

```

## Data Cleaning and Description

We select part of the variables from the original dataset, add a few variables and change their names to make them easy to read.

#### NYC Crime data

  * `date`. Date the crime occurs.
  * `hour_of_day`. Occurring time of the crimes
  * `borough`. Borough the crime occurs.
  * `level`. Level of offense: felony, misdemeanor, violation.
  * `offense`. Description of offense.
  * `offense_classification`. Description of internal classification of offense.
  * `place`. Specific description of premises; grocery store, residence,street, etc.
  * `suspect_age`. Suspect’s Age Group.
  * `suspect_race`. Suspect’s Race Description.
  * `suspect_sex`. Suspect’s Sex Description.
  * `victim_age`. Victim’s Age Group.
  * `victim_race`. Victim’s Race Description.
  * `victim_sex`. Victim’s Sex Description.
  * `day_of_week`. Day of week that a crime occurs.
  * `mon`. Months in a year that a crime occurs.
  * `latitude`. Latitude of the place the crime occurs.
  * `longitude`. Longitude of the place the crime occurs.

```{r}
weather = rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00094789", "USW00014732"),
    var = c("PRCP", "SNOW", "SNWD", "TMAX", "TMIN"),
    date_min = "2016-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00094789 = "JFK Airport_NY",
      USW00014732 = "La Guardia_NY"),
    
    
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

average_weather <- weather %>% 
  group_by(date) %>%
  summarise(
    mean_snow = mean(snwd),
    mean_prcp = mean(prcp),
    mean_tmin = mean(tmin),
    mean_tmax = mean(tmax)
  ) 

```

#### Weather Data

I selected weather data collected from 2016-01-01 to 2017-12-31 and three weather stations in New York. I 
calculated mean weather data across the three weather stations.

  * `mean_snow`. snow depth(mm).
  * `mean_prcp`. precipitation(tenths of mm)
  * `mean_tmin`. minimum temperature of a day
  * `mean_tmax`. maximum temperature of a day






 

  
