---
title: "Crime map"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    theme: journal
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(sf) #for analyzing spatial data
library(rgdal) #note that `rgdal` requires another package `sp`; for importing zips shapefile, transforming CRS, and transforming coordinates into a SpatialPointsDataFrame
library(leaflet) #for interactive maps
library(tigris) #geojoin

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .2,
  out.width = "90%",
  dpi = 300,
  warning = FALSE,
  message = FALSE
)
```

We first defined our dataframe `crime_df_map`.

```{r, message=FALSE, warning=FALSE}
#read crime_df data
crime_df = read_csv("data/crime_df.csv") %>%
  janitor::clean_names() %>%
  rename(case_number = x1)

#define crime_df_map for our map
crime_df_map = drop_na(crime_df, latitude)
```

# Simple map

Get a first look of our dataframe by making a simple map draft. Drag the slider below to see spatial distribution of crime incidents in different time period during a day.

```{r, message=FALSE}
crime_plot =
  crime_df %>% 
  filter(!is.na(borough)) %>%
  plot_ly(
    lat = ~latitude, 
    lon = ~longitude, 
    frame = ~hour_of_day,
    type = "scattermapbox", 
    mode = "markers", 
    alpha = 0.2,
    color = ~borough) %>% 
  layout(
    title = "Spatial distribution of crime incidents",
    mapbox = list(
      style = 'carto-positron',
      zoom = 9,
      center = list(lon = -73.9, lat = 40.7)),
    legend = list(title=list(text='<b> Borough </b>'))
    ) %>%
  animation_slider(
    currentvalue = list(prefix = "Hour of day: ", font = list(color="black"))
  )
crime_plot
```

# Crime rate interactive map categorized by modified zipcodes

We want to draw a crime rate map to examine the crime rate in each zip code region. Crime rate for a zip code region is defined as number of crime incidents occurred in the zip code region during 2016-01-01 and 2017-12-31, divided by population size in the corresponding zip code region.

## Data preparation

We first imported zips shapefile and transform CRS.

```{r}
zips = readOGR("data/cb_2015_us_zcta510_500k/cb_2015_us_zcta510_500k.shp", verbose=FALSE)
zips = spTransform(zips, CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
```
Then we matched each crime incident to a zip code region by utilizing their coordinates variables `longitude` and `latitude`. We included a `zip` variable to `crime_df_map` data frame, representing corresponding zip codes of each crime incident.

```{r}
#extract only lon and lat
crime_lat_long = select(crime_df_map, longitude, latitude)

#transform coordinates into a SpatialPointsDataFrame
crime_spdf = SpatialPointsDataFrame(coords = crime_lat_long, data = crime_lat_long, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

#subset only the zipcodes in which points are found
crime_zips = zips[crime_spdf, ]

#redefine crime_df_map to include zipcodes
crime_df_map = cbind(crime_df_map, over(crime_spdf, crime_zips[,"ZCTA5CE10"]))
crime_df_map = rename(crime_df_map, zip = ZCTA5CE10)
```

`r sum(is.na(crime_df_map$zip))` crime incidents can't be matched to a zip code, so we drop them.

```{r}
crime_df_map =
  crime_df_map %>%
  drop_na(zip)
```

In order to make meaningful data analysis, we want zip codes to represent regions with similar population. However, some zip code regions are very small, among which there are even some buildings that have their own zip codes. To deal with the challenges of zip codes, we convert zip codes to modified zip codes to account for discrepancies in population size. Modified zip codes represent regions with similar population size. In order to make the conversion, we use a conversion table `ZCTA-to-MODZCTA.csv` obtained from [here]( https://github.com/nychealth/coronavirus-data/blob/master/Geography-resources/ZCTA-to-MODZCTA.csv).

We first read the conversion table `ZCTA-to-MODZCTA.csv`.

```{r}
zcta_conv = read_csv("data/ZCTA-to-MODZCTA.csv")
zcta_conv$ZCTA = as.character(zcta_conv$ZCTA)
zcta_conv$MODZCTA = as.character(zcta_conv$MODZCTA)
```

We matched `zip` in `crime_df_map` with corresponding modified zip code, represented by a new variable `mod_zip`.

```{r}
crime_df_map = crime_df_map %>%
  left_join(rename(zcta_conv, zip = ZCTA), by = "zip") %>%
  rename(mod_zip = MODZCTA)
```

`r sum(is.na(crime_df_map$mod_zip))` crime incidents can't be matched to a modified zip code and leave NA values in the `mod_zip` column. This is because the `r sum(is.na(crime_df_map$mod_zip))` crime incidents happened in zip code 10550 or 10704, which are not in New York City. Therefore, we dropped the `r sum(is.na(crime_df_map$mod_zip))` observations.

```{r}
crime_df_map = drop_na(crime_df_map, mod_zip)
```

In order to make an interactive map based on modified zip codes, we use `modzcta` shapefile obtained from [here](https://github.com/nychealth/coronavirus-data/tree/master/Geography-resources). Note this shapefile is based on modified zip codes instead of standard zip codes.

```{r}
modzcta = st_read("data/MODZCTA_2010/MODZCTA_2010.shp", quiet=TRUE) %>%
  janitor::clean_names()
```

We included shapefile information into `crime_df_map`.

```{r, warning=FALSE}
crime_df_geo =
  geo_join(modzcta, crime_df_map, "modzcta", "mod_zip", how = "inner") %>%
  rename(mod_zip = modzcta)
```

We can then obtain number of crimes in each zip region, sorting in descending order.

```{r}
number_of_crimes = 
  crime_df_geo %>% 
  group_by(mod_zip) %>% 
  summarize(number_of_crimes = n()) %>%
  arrange(desc(number_of_crimes))
```

In order to obtain crime rate, we got population of each zip code region from [here](https://worldpopulationreview.com/zips/new-york).

```{r}
pop_zip = read_csv("data/csvData.csv") %>%
  janitor::clean_names()
pop_zip$zip = as.character(pop_zip$zip)

#filter out regions in New York City
pop_zip = pop_zip %>%
  filter(county %in% c("New York", "Kings", "Queens", "Bronx", "Richmond"))
```

The dataset `pop_zip` lacks population data in zip code region 10065, so we manually added it in ([data source here](https://censusreporter.org/profiles/86000US10065-10065/)).

```{r}
pop_zip[nrow(pop_zip) + 1,] = list("10065","	
New York City", "New York", 30339)
```

We then converted zip code to modified zip code, and got the population in each modified zip code region.

```{r}
pop_mod_zip = pop_zip %>%
  left_join(rename(zcta_conv, zip = ZCTA), by = "zip") %>%
  rename(mod_zip = MODZCTA) %>%
  group_by(mod_zip) %>%
  summarize(population = sum(population))
```

We added population of each modified zip code region into `number_of_crimes`, and calculated crime rate in each modified zip code region. Crime rates were sorted in descending order.

```{r}
crime_rate = left_join(number_of_crimes, pop_mod_zip, by = "mod_zip")
crime_rate$crime_rate = crime_rate$number_of_crimes / crime_rate$population * 100
crime_rate = 
  crime_rate %>%
  select(mod_zip, crime_rate) %>%
  group_by(crime_rate) %>%
  arrange(desc(crime_rate))
```

## Interactive map of crime rate.

```{r}
#label
labels = sprintf(
  "<strong>%s</strong><br/>Crime rate is %g&percnt;<br/>during the period from 2016-01-01 to 2017-12-31.", crime_rate$mod_zip, crime_rate$crime_rate) %>%
  lapply(htmltools::HTML)

#color palette
pal = colorBin(palette = "OrRd", 9, domain = crime_rate$crime_rate)

map_interactive = crime_rate %>%
  st_transform(crs = "+init=epsg:4326") %>%
  leaflet() %>%
  addProviderTiles(provider = "CartoDB.Positron") %>%
  addPolygons(label = labels,
              stroke = FALSE,
              smoothFactor = 0.5,
              opacity = 1,
              fillOpacity = 0.7,
              fillColor = ~ pal(crime_rate),
              highlightOptions = highlightOptions(weight = 5,
                                                  fillOpacity = 1,
                                                  color = "black",
                                                  opacity = 1,
                                                  bringToFront = TRUE)) %>%
  addLegend("bottomright",
            pal = pal,
            values = ~ crime_rate,
            title = "Crime cases per 100 residents<br/>during the period from<br/>2016-01-01 to 2017-12-31",
            opacity = 0.7) %>%
  addTiles("Interactive map of crime rate")
#saveWidget(map_interactive, "nyc_crime_rate_map.html")
map_interactive
```

# Crime data Shiny map categorized by modified zip codes and 4 weeks interval

## Data preparation

We cut `date` into 4 weeks interval, and want to visualize crime counts and crime rate in each modified zip code region for every 4 weeks since 2016-01-01 by using R Shiny.

```{r}
crime_month = crime_df_map %>% #use crime_df_map instead of crime_df_geo here because crime_df_geo is too large 
  group_by(mod_zip, date) %>%
  summarize(crime_counts = n())
crime_month$date = cut(crime_month$date, "28 days")
crime_month =  crime_month %>%
  group_by(mod_zip, date) %>%
  summarize(crime_counts = sum(crime_counts)) %>%
  rename(month_following = date) %>%
  filter(month_following != "2017-12-29")  #observations for 2017-12-29 removed for too few data
  
#add back mod_zip, crime_rate, and geometry
crime_month = left_join(crime_month, pop_mod_zip, by = "mod_zip") %>%
  select(mod_zip, population, month_following, crime_counts)
crime_month$crime_rate = crime_month$crime_counts / crime_month$population
crime_month = crime_month %>%
  select(month_following, mod_zip, population, crime_counts, crime_rate) %>%
  arrange(month_following)
crime_month = geo_join(modzcta, crime_month, "modzcta", "mod_zip", how = "inner")
crime_month$month_following = as.Date(crime_month$month_following, format = "%Y-%m-%d")
```

```{r}
saveRDS(crime_month, "r_shiny_map/crime_month.RDS") #save for R Shiny
```

## Shiny app

The Shiny app can be accessed [here](https://liyc5215.shinyapps.io/r_shiny_map/).