---
title: "Modeling"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: journal
---

In some articles, we knew that weather affect crime rate. 
For example, the hotter the weather is, the more likely people are to commit crimes. 
The worse the weather is, the less likely people are to go out and commit crimes. 
Here we want to predict number of crimes each day in New York as our outcome and precipitation, 
snow depth, maximum temperature and minimum temperature as predictors.


```{r, include=FALSE}
library(tidyverse)
library(lubridate)
library(modelr)
library(ggplot2)
library(viridis)
library(car)
knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  echo = T,
  warning = F,
  cache = F
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis",
  digits = 3
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
set.seed(1)
```



```{r include = FALSE}
crime_df <- read.csv("data/crime_df.csv")

weather <- read.csv("data/weather.csv")

new_crime_df = crime_df %>%
  group_by(date,mon,day_of_week)%>%
  summarize(
    total_crime = n(),
    ) 


crime_weather = merge(new_crime_df, weather, by = "date")
crime_weather <- crime_weather %>%
  mutate(
    month = factor(mon)
  )
```

First, let's check the distribution of number of crime cases each day in New York 
```{r}
ggplot(data = crime_weather,aes(x = total_crime)) + geom_histogram()
```
From the distribution of daily cases, we noticed that it is approximately normally 
distributed, therefore we don't need to transform the value of the outcome.

First we use scatter plot to investigate the association between weather and cases
and crime 


```{r}
ggplot(crime_weather, aes(x = mean_tmax, y = total_crime)) + geom_point() 
ggplot(crime_weather, aes(x = mean_tmin, y = total_crime)) + geom_point() 
ggplot(crime_weather, aes(x = mean_prcp, y = total_crime)) + geom_point() 
ggplot(crime_weather, aes(x = mean_snow, y = total_crime)) + geom_point()
```
 
From the plots, we noticed that all of the four predictors may have association 
with the number of cases of crime in New York, So I proposed the following models. 
We also want to investigate whether the month and day of week can be predictors.

## Fitting Model
 
Firstly, I use month, weekdays and all weather data as predictors
```{r}
full_model = lm(total_crime ~ month + day_of_week + mean_tmin + mean_tmax + mean_prcp + mean_snow, data = crime_weather)
summary(full_model) %>%
  broom::tidy() %>%
  knitr::kable(digits = 2)

```

I noticed that both tmax and tmin are not significant, which may be attributed to their
correlation, so I calculated the VIF. 

```{r}
vif(full_model) %>%
  knitr::kable()
```

It is obvious that the VIF of tmin and tmax is high, so I chose to drop tmin. Besides,
I found that some months are not significantly different from the reference category.
So, what if we drop the month?

The second proposed model is total_crime ~  day_of_week + mean_tmax + mean_prcp + mean_snow
```{r}
model1 = lm(total_crime ~  day_of_week + mean_tmax + mean_prcp + mean_snow, data = crime_weather)
summary(model1) %>%
  broom::tidy() %>%
  knitr::kable(digits = 2)
```
This time, the tmax becomes significant. And it has a positive effect on crime rate, 
which validate our assumption. Besides, all other variables are significant. 


But we noticed that some months are significantly different from reference catergory, so we want to compare model1 and model with month.
```{r}
model2 = lm(total_crime ~ month +day_of_week +  mean_tmax + mean_prcp + mean_snow, data = crime_weather)
summary(model2) %>%
  broom::tidy() %>%
  knitr::kable(digits = 2)

anova(model1, model2)
```
We use ANOVA to compare the two models. Since p-value is smaller than 0.05, we should include month in our model.

We also want to try to include a interaction term in our model.
```{r}
model3 <- lm(total_crime ~  month +day_of_week +  mean_tmax + mean_prcp + mean_snow + mean_tmax*day_of_week, data = crime_weather )
summary(model3) %>%
    broom::tidy() %>%
  knitr::kable(digits = 2)
```
In model 3 , we assume that there is an interaction between temperature and day of week. However, some of the interaction term are not significant.

We also want to know whether to include months and weekdays in our model.
```{r}
model4 <- lm(total_crime ~ mean_tmax + mean_prcp + mean_snow, data = crime_weather)
summary(model4) %>%
  broom::tidy() %>%
  knitr::kable(digits = 2)
```

## Cross Validation
We want to decide which model is the best to predict the daily number of crime cases in New York. We use cross validation to compare the models.
```{r}
cv_df <- 
  crossv_mc(crime_weather, 100) %>%  
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r}
cv_df <- 
  cv_df %>% 
  mutate(
    model1 = map( train, ~lm(total_crime ~ month + day_of_week + mean_tmax + mean_prcp + mean_snow, data = crime_weather)),
    model2 =  map( train, ~lm(total_crime ~ day_of_week +  mean_tmax + mean_prcp+mean_snow, data = crime_weather)),
    model3 =  map( train, ~lm(total_crime ~ month +day_of_week +  mean_tmax + mean_prcp + mean_snow +mean_tmax*day_of_week, data = crime_weather)),
    model4 = map( train, ~lm( total_crime ~ mean_tmax + mean_prcp + mean_snow, data = crime_weather))) %>%
   mutate(
    rmse_model1 = map2_dbl( model1,  test, ~rmse(model = .x, data =.y)),
    rmse_model2 = map2_dbl( model2,  test, ~rmse(model = .x, data =.y)),
    rmse_model3 = map2_dbl( model3,  test, ~rmse(model = .x, data =.y)),
    rmse_model4 = map2_dbl( model4, test, ~rmse(model = .x, data =.y))
  ) 
```


```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

cv_df %>% 
  summarize(
    rmse_1 = mean(rmse_model1),
    rmse_2 = mean(rmse_model2),
    rmse_3 = mean(rmse_model3),
    rmse_4 = mean(rmse_model4)
  ) %>%
  knitr::kable()
```
## Model Diagnostics

```{r}
par(mfrow = c(2,2))
plot(model3)
```



 Model3's root means square error is the smallest in the four models, which may indicates it is the best model in the four models. We then did a model diagnostic to check if the model assumptions. Although there are some outliers as shown in the plots, the Q-Q plot, residuals VS fitted plot and Scale-Location plot indicates the residuals are normally distributed and variance of residuals is constant. Therefore, model3 could be useful for predicting the total number of crime cases each day in New York, which means month, day of week, temperature, precipitation , snow depth and a interaction term between day of week and temperature are all useful predictors if we want to predict the number of crime cases on a specific day in New York.

