---
title: "Statistical test"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: journal
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  echo = T,
  warning = F,
  cache = F
)

crime_df <- read.csv("data/crime_df.csv")
weather_df <- read.csv("data/weather.csv")
```

## Chi-squared test on year and number of cases each month.

We want to verify whether the number distribution of umber of crime cases of each 
month is  related to the year. Therefore we performed Chi-squared test for 
comparison of year of 2016 and 2017 :

$H_0$ : The number distribution of  number of crime cases of each month was the same in 2016 and 2017

$H_1$ : The number distribution of  number of crime cases of each month was different in 2016 and 2017


```{r}
crime_weather <- 
  crime_df %>%
  group_by(date)%>%
  summarise(total_crime = n())%>%
  merge(weather_df,by = "date")

df_1 <- crime_df %>%
  mutate (
    year = year(date)
  )

table(df_1$year, df_1 $mon)%>%
  knitr::kable(digits = 3)
```

```{r}
chisq.test(table(df_1$year, df_1 $mon))
```
As the result told us, the p-value < 0.05 , therefore we reject the null 
hypothesis and conclude that the distribution of number of crimes cases
of each month is significantly different between 2016 and 2017.


## ANOVA between number of crimes per day and weather

We also concerned about whether weather conditions will affect New York City's 
crime frequency per day. Therefore it is important to test hypotheses according 
to different weather conditions and crime cases per day.

At first, we divided the daily weather into four categories according to whether 
the precipitation and rainfall are greater than 0.
```{r}
crime_weather_2 <-  crime_weather %>%
  mutate(weather = case_when(mean_prcp == 0 & mean_snow == 0 ~ 'not_rainy_or_snowy',
                           mean_prcp >0 & mean_snow >0 ~ 'rainy_and_snowy',
                           mean_prcp == 0 & mean_snow >0 ~ 'snowy',
                           mean_prcp >0 & mean_snow == 0 ~ "rainy")) %>%
  select(date, total_crime,  weather)

crime_weather_2 %>%
  group_by(weather)%>%
  summarise(mean_daily_crime = mean(total_crime),
  Standard_deviation = sd(total_crime)) %>%
  knitr::kable(digits = 3)


```


We then conducted one-way ANOVA to test if the true mean daily crime cases of 
different weather is the same.

$H_0$: There is no difference in the average number of crimes per day under 
different weather conditions in New York.

$H_1$: At least the average number of crimes per day under two different weather 
conditions are different.

First, Let's create box-plot of our data.

```{r}
ggplot(data = crime_weather_2, aes(x = weather, y = total_crime, fill = weather)) + geom_boxplot()
```

From the plot, the distribution of daily crime number seems different in 
different weather.

And we conducted a ANOVA test
```{r}
res.aov <- aov(total_crime ~ weather, data = crime_weather_2)
summary(res.aov) 
```

From the result, the p-value is smaller than 0.05,  therefore we concluded that 
the average number of crimes per day is correlated to weather. 

```{r}
crime_weather_3 = crime_df %>%
  merge(weather_df , by = "date")%>%
  mutate(weather = case_when(mean_prcp == 0 & mean_snow == 0 ~ 'not_rainy_or_snowy',
                           mean_prcp >0 & mean_snow >0 ~ 'rainy_and_snowy',
                           mean_prcp == 0 & mean_snow >0 ~ 'snowy',
                           mean_prcp >0 & mean_snow == 0 ~ "rainy")) 

df_crime_weather <- crime_weather_3 %>%
  group_by(weather, offense) %>%
  summarise(Total_number_of_offense = n()) %>%
  pivot_wider(
    names_from = "weather",
    values_from = "Total_number_of_offense"
  ) 
df_crime_weather[is.na(df_crime_weather)] = 0
df_crime_weather %>%
  head(20) %>%
  knitr::kable(digits = 3)
```

## Chi-squared test on weather and number of different types of offense

We also want to know whether the distribution of number of different type of 
crime cases across different weather are the same. Therefore, we performed 
another Chi-squared test on number of different type of crimes and weather.


$H_0$ : The number distribution of each type of crime was the same in different weathers.

$H_1$ : The number distribution of each type of crime was different in different weathers.

```{r}
chisq.test(table(crime_weather_3$weather, crime_weather_3$offense))
```
We can see p-value is smaller than 0.05, so we reject the null hypothesis and 
conclude that there's at least one type of weather's proportions of number of 
different type of crimes is different from others at 0.05 significant level.  
