---
title: "Statistical test"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(
  fig.height = 6,
  fig.width = 8,
  message = F,
  echo = T,
  warning = F,
  cache = F
)

crime_df <- read.csv("data/crime_df.csv")
weather_df <- read.csv("data/weather.csv")
```

## Chi-squared test on year and types of offense

We want to verify whether the number distribution of different crime types is 
related to the year. Therefore we performed Chi-squared test for comparison of 
year of 2016 and 2017 :

$H_0$ : The number distribution of each type of crime was the same in 2016 and 2017

$H_1$ : The number distribution of each type of crime was different in 2016 and 2017


```{r}
crime_weather <- 
  crime_df %>%
  group_by(date)%>%
  summarise(total_crime = n())%>%
  merge(weather_df,by = "date")

df_1 <- crime_df %>%
  mutate (
    year = year(date)
  )%>%
  group_by(year,offense) %>%
  summarise(total_by_level = n()) %>%
  pivot_wider(
    names_from = "year",
    values_from = "total_by_level"
  )%>%
  na.omit()
df_1 %>%
  head() %>%
  knitr::kable(digits = 3)
```
```{r}
chisq.test(table(df_1$`2016`, df_1 $`2017`))
```
As the result told us, the p-value < 0.05 , therefore we reject the null 
hypothesis and conclude that the distribution of number of different type of 
crime cases are different in 2016 and 2017.

## Chi-squared test on Borough and number of different types of offense

We also want to know whether the distribution of number of different type of 
crime cases across different borough are the same. Therefore, we performed 
another Chi-squared test on number of different type of crimes and borough.
Because the population of Staten Island is small, so we did not take it into 
consideration.

$H_0$ : The number distribution of each type of crime was the same in different boroughs.

$H_1$ : The number distribution of each type of crime was different in different boroughs.



```{r}
crime_df_3 <- crime_df %>%
  group_by(borough, offense)%>%
  summarise(
    number = n()
  ) %>%
  filter(borough %in% c("BRONX", "BROOKLYN","MANHATTAN", "QUEENS")) %>%
  pivot_wider(
    names_from = "borough",
    values_from = "number"
  )
crime_df_3%>%
  head()%>%
  knitr::kable(digits = 3)
```
```{r}
chisq.test(crime_df_3$BRONX,crime_df_3$BROOKLYN, crime_df_3$MANHATTAN, crime_df_3$QUEENS)
```
We can see p-value is smaller than 0.05, so we reject the null hypothesis and 
conclude that there's at least one borough's proportions of violation counts 
for week days is different from others at 0.05 significant level.  


## ANOVA between number of crimes per day and weather

We also concerned about whether weather conditions will affect New York City's 
crime frequency per day. Therefore it is important to test hypotheses according 
to different weather conditions and crime cases per day.

At first, we divided the daily weather into four categories according to whether 
the precipitation and rainfall are greater than 0.
```{r}
crime_weather_2 <-  crime_weather %>%
  mutate(weather = case_when(mean_prcp == 0 & mean_snow == 0 ~ 'not_rainy_or_snowy',
                           mean_prcp >0 & mean_snow >0 ~ 'rainy_and_snowy',
                           mean_prcp == 0 & mean_snow >0 ~ 'snowy',
                           mean_prcp >0 & mean_snow == 0 ~ "rainy")) %>%
  select(total_crime, weather)
crime_weather_2 %>%
  head()%>%
  knitr::kable(digits = 3)
```

We then conducted one-way ANOVA to test if the true mean daily crime cases of 
different weather is the same.

$H_0$: There is no difference in the average number of crimes per day under 
different weather conditions in New York.

$H_1$: At least the average number of crimes per day under two different weather 
conditions are different.

First, Let's create box-plot of our data.

```{r}
ggplot(data = crime_weather_2, aes(x = weather, y = total_crime, fill = weather)) + geom_boxplot()
```

From the plot, the distribution of daily crime number seems different in 
different weather.

And we conducted a ANOVA test
```{r}
res.aov <- aov(total_crime ~ weather, data = crime_weather_2)
summary(res.aov) 
```

From the result, the p-value is smaller than 0.05,  therefore we concluded that 
the average number of crimes per day is correlated to weather. 

